
\section{IceStorm (iCE40)}
IceStorm \cite{project-icestorm} is used for iCE40 bit-stream creation. It relies on \textbf{yosys} for synthesis and \textbf{NextPNR} for place and route.

Supports all package variants of LP1K, LP4K, LP8K and HX1K, HX4K, HX8K, LP384 and UltraPlus devices.

Does not support iCE40 LM, Ultra and UltraLite
\section{Trellis (ECP5)}
Project Trellis enables a fully open-source flow for ECP5 FPGAs using \textbf{Yosys} for Verilog synthesis and \textbf{nextpnr} for place and route. Project Trellis itself provides the device database and tools for bit-stream creation.

The following features are currently working in the Yosys -> nextpnr -> Trellis flow.

\begin{itemize}
\item Logic slice functionality, including carries
\item Distributed RAM inside logic slices
\item All internal interconnect
\item Basic IO, including tristate, using TRELLIS\_IO primitives. LPF files and DDR inputs/outputs
\item Block RAM, using either inference in Yosys or manual instantiation of the DP16KD primitive
\item Multipliers using manual instantiation of the MULT18X18D primitive. 
\newline Inference and more advanced DSP features are not yet supported.
\item Global networks (automatically promoted and routed in nextpnr)
\item PLLs
\item Transcievers (DCUs)
\end{itemize}

\section{NextPNR} 
    
NextPNR is a portable FPGA place and route tool that aims to be a vendor neutral, timing driven, FOSS FPGA place and route tool.

Currently NextPNR supports:
\begin{itemize}
\item Lattice iCE40 devices through project IceStorm
\item Lattice ECP5 devices through project Trellis, currently experimental
\item "generic" back-end for user-defined architectures, currently experimental
\item Xilinx 7 Series could be supported in the future through project X-Ray 
\end{itemize}

\section{X-Ray (Xilinx 7)}
Project X-Ray documents the Xilinx 7-Series FPGA architecture to enable development of open-source tools.



\section{Symbiflow}
Symbiflow aims to be an open source flow for generating bit-streams from Verilog. 

Since it uses Yosis, icestorm, trellis, xray and nextpnr for each of the supported platforms, device support is at the same level as the underlying projects.


\section{IceStudio}
IceStudio is a visual editor for open FPGA boards. Built on top of the Icestorm project using Apio.

It implements Graphic design -> Verilog, PCF -> Bistream -> FPGA

It supports the following devices:

\begin{itemize}
\item HX1K
\item HX8K
\item LP8K
\item UP5K
\end{itemize}

It is the easiest to setup as it handles all the tool-chain and device drivers downloads and it comes as an AppImage.

\section{APIO}
APIO is an Open source ecosystem for open FPGA boards inspired by PlatformIO
It is built on top of Icestorm project and currently supports the ICE40 FPGA family
A list supported boards is available on their github page, these include the following ICE40 families:

\begin{itemize}
\item HX1K
\item HX8K
\item LP8K
\item UP5K
\end{itemize}

Documentation page: \url{http://apiodoc.readthedocs.io/}

GitHub Repo: \url{https://github.com/FPGAwars/apio}

APIO comes with a set of packages:

\begin{itemize}
\item Icestorm for iCE40 FPGA synthesis, place \& route and configuration tools
\item Simulation viewer. GTKWave project (only for Windows)
\item Verilog simulation and synthesis tool. Icarus Verilog project
\item Verilog HDL simulator. Verilator project
\end{itemize}

APIO also has an Atom editor package called apio-ide. Most of APIO functionality is available from related Atom menu items

\begin{itemize}
\item Verify
\item Build
\item Upload
\item Simulation
\item Time analysis
\end{itemize}


\section{Intel Quartus Prime Lite Edition and ModelSim - Intel FPGA Starter Edition}
Intel Quartus Prime is programmable logic device design software produced by Intel; prior to Intel's acquisition of Altera the tool was called Altera Quartus II. Quartus Prime enables analysis and synthesis of HDL designs, which enables the developer to compile their designs, perform timing analysis, examine RTL diagrams, simulate a design's reaction to different stimuli, and configure the target device with the programmer. Quartus Prime includes an implementation of VHDL and Verilog for hardware description, visual editing of logic circuits, and vector waveform simulation.\cite{Wikipedia_Quartus}

\section{CubicBoard}
Cubicboard claims to be an open-source FPGA project for the Cyclone-V family.
They do provide Open Hardware under Apache 2.0 license although source files are for Protel/Altium EDA.
The provided software is a VirtualBox image of Ubuntu having Intel Quartus preinstalled.

\section{Vivado}
This is the official software from Xilinx. The HL WEBPack edition does not require a license and there is a Linux version.

Vivado Design Suite is a software suite produced by Xilinx for synthesis and analysis of HDL designs, superseding Xilinx ISE with additional features for system on a chip development and high-level synthesis. Vivado represents a ground-up rewrite and re-thinking of the entire design flow (compared to ISE), and has been described by reviewers as "well conceived, tightly integrated, blazing fast, scalable, maintainable, and intuitive"\cite{Wikipedia_Vivado}

% \idea